---
title: "DV_RProject3"
output: html_document
---

***Members***
Raymundo Cortez
Yuyin Wang
Chigozi Yejekwe

###R_ETL
```
require(tidyr)
require(dplyr)
require(ggplot2)
setwd("~/DataVisualization/DV_RProject3/01 Data")
file_path <- "Salary_Information_for_Industrial_Development_Agencies.csv"

df <- read.csv(file_path, stringsAsFactors = FALSE)
# Replace "." (i.e., period) with "_" in the column names.
names(df) <- gsub("\\.+", "_", names(df))
#shorten names longer than 30, or reserved words in SQL
df <- rename(df,Paid_by_State_or_Local_Gov=Paid_by_State_or_Local_Government)
df <- rename(df,Grp = Group)

measures <- c("Base_Annualized_Salary","Actual_Salary_Paid","Overtime_Paid","Performance_Bonus","Extra_Pay","Other_Compensation","Total_Compensation")

# Get rid of special characters in each column.
for(n in names(df)) {
  df[n] <- data.frame(lapply(df[n], gsub, pattern="[^ -~]",replacement= ""))
}

dimensions <- setdiff(names(df), measures)
if( length(measures) > 1 || ! is.na(dimensions)) {
  for(d in dimensions) {
    # Get rid of " and ' in dimensions.
    df[d] <- data.frame(lapply(df[d], gsub, pattern="[\"']",replacement= ""))
    # Change & to and in dimensions.
    df[d] <- data.frame(lapply(df[d], gsub, pattern="&",replacement= " and "))
    # Change : to ; in dimensions.
    df[d] <- data.frame(lapply(df[d], gsub, pattern=":",replacement= ";"))
  }
}


# Get rid of all characters in measures except for numbers, the - sign, and period.dimensions
if( length(measures) > 1 || ! is.na(measures)) {
  for(m in measures) {
    df[m] <- data.frame(lapply(df[m], gsub, pattern="[^--.0-9]",replacement= ""))
  }
}

write.csv(df, paste(gsub(".csv", "", file_path), ".reformatted.csv", sep=""), row.names=FALSE, na = "")

tableName <- gsub(" +", "_", gsub("[^A-z, 0-9, ]", "", gsub(".csv", "", file_path)))
sql <- paste("CREATE TABLE", tableName, "(\n-- Change table_name to the table name you want.\n")
if( length(measures) > 1 || ! is.na(dimensions)) {
  for(d in dimensions) {
    sql <- paste(sql, paste(d, "varchar2(4000),\n"))
  }
}
if( length(measures) > 1 || ! is.na(measures)) {
  for(m in measures) {
    if(m != tail(measures, n=1)) sql <- paste(sql, paste(m, "number(38,4),\n"))
    else sql <- paste(sql, paste(m, "number(38,4)\n"))
  }
}
sql <- paste(sql, ");")
cat(sql)

#repeat for second csv
file_path <- "Actual_avg_loc_salary.csv"

df <- read.csv(file_path, stringsAsFactors = FALSE)

# Replace "." (i.e., period) with "_" in the column names.
names(df) <- gsub("\\.+", "_", names(df))

df <- rename(df,Grp = Group)

measures <- c("AVG_LOC_BASE_ANU_SALARY","AVG_LOC_ACT_SALARY","AVG_LOC_OVERTIME","AVG_LOC_BONUS","AVG_LOC_EXTRA_PAY","AVG_LOC_TOT_COMP","AVG_LOC_OTHER_COMP")

# Get rid of special characters in each column.
for(n in names(df)) {
  df[n] <- data.frame(lapply(df[n], gsub, pattern="[^ -~]",replacement= ""))
}

dimensions <- setdiff(names(df), measures)
if( length(measures) > 1 || ! is.na(dimensions)) {
  for(d in dimensions) {
    # Get rid of " and ' in dimensions.
    df[d] <- data.frame(lapply(df[d], gsub, pattern="[\"']",replacement= ""))
    # Change & to and in dimensions.
    df[d] <- data.frame(lapply(df[d], gsub, pattern="&",replacement= " and "))
    # Change : to ; in dimensions.
    df[d] <- data.frame(lapply(df[d], gsub, pattern=":",replacement= ";"))
  }
}


# Get rid of all characters in measures except for numbers, the - sign, and period.dimensions
if( length(measures) > 1 || ! is.na(measures)) {
  for(m in measures) {
    df[m] <- data.frame(lapply(df[m], gsub, pattern="[^--.0-9]",replacement= ""))
  }
}

write.csv(df, paste(gsub(".csv", "", file_path), ".reformatted.csv", sep=""), row.names=FALSE, na = "")

tableName <- gsub(" +", "_", gsub("[^A-z, 0-9, ]", "", gsub(".csv", "", file_path)))
sql <- paste("CREATE TABLE", tableName, "(\n-- Change table_name to the table name you want.\n")
if( length(measures) > 1 || ! is.na(dimensions)) {
  for(d in dimensions) {
    sql <- paste(sql, paste(d, "varchar2(4000),\n"))
  }
}
if( length(measures) > 1 || ! is.na(measures)) {
  for(m in measures) {
    if(m != tail(measures, n=1)) sql <- paste(sql, paste(m, "number(38,4),\n"))
    else sql <- paste(sql, paste(m, "number(38,4)\n"))
  }
}
sql <- paste(sql, ");")
cat(sql)
```

###Summary
The first of our datasets is the "Salary Information for Industrial Development Agencies". The Dataset includes information on employees for Industrial Development Agencies in the State of New York. The information includes for each employee, Title, Group (Executive, Managerial, etc...), Department, Pay Type, Exempt Indicator, Paid by state or local government, paid by another entity, base annualized salary, actual salary paid, overtime paid, performance bonus, extra pay, other compensation and total compensation.
The second dataset is taken from the "Salary Information for Local Development Corporations". The Dataset included information on employees for Local Development Corporations in the State of New York. The averages of their base annualized salary, actual salary paid, overtime paid, performance bonus extra pay, other compensation and total compensation based on Group were then compiled into a Dataset named "Actual Average Local Salary".

###Workflows and graphs

***Graph 1***
The Following Graph compares the differences of Industrial Agencies' Employee's total compensation from the average of Local Companies' Employee's total compensation. The graph then displays a boxplot for each group showing the range of differences. If the values are below zero, then they are being paid less than the average Local Company Employee, if greater than zero they are being paid more. After observing the graph it appears that most Industrial Employees get paid less than the average Local Corporation Employee in all groups save for Executives.

```{r, message=FALSE, warning=FALSE}

source("~/DataVisualization/DV_RProject3/03 Visualizations/Workflow1.R", echo = TRUE)
```

***Graph 2***
Average Overtime Comparison of NY Development Agencies

It appears that local companies (commercially created) log a lot of overtime, especially Operational or Administrative/Clerical type jobs. Executive and Techincal/Engineering type jobs appears to have essentially no overtime in both industrial agencies (government created) and local companies. Overall, local company employees record a siginificantly more amount of overtime than their industrial agency counterparts.

```{r, message=FALSE, warning=FALSE}

source("~/DataVisualization/DV_RProject3/03 Visualizations/Workflow2.R", echo = TRUE)
```

***Graph 3***
Comparing Salary Paid by Group and Name, to test the idea of income equality between genders.

```{r, message=FALSE, warning=FALSE}

source("~/DataVisualization/DV_RProject3/03 Visualizations/Workflow3.R", echo = TRUE)
```